---
title: "Activity Monitoring"
output: html_document
---


```{r, echo=FALSE}
#File is downloaded into workspace.  
  
setwd("~/Documents/AngelaStuff/5-ReproducibleResearch/PeerAssesment1")
#fileURL<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
#download.file(fileURL, destfile="activity.csv", method="curl")
#dateDownloaded<-date()
library(lattice)
library(plyr)
#Read "activity.csv" as csv data file.
d<-read.csv("activity.csv", header = TRUE, quote="\"", stringsAsFactors=FALSE)
```
  
    
#What is mean total number of steps taken per day?   
```{r stepsPerDay, echo=FALSE}
#Get the average number of steps taken each day and  
sumStepsPerday<-tapply(d$steps, as.vector(d$date), sum)
#Plot a histogram
hist(sumStepsPerday, breaks = 20, ylab="Number of Days",
    xlab="Total Number of Steps",
    main="Histogram of Number of Steps Taken Each Day
    Between Oct-01-2012 and Nov-30-2012")
#-calculate the median and mean number of steps      
# for all the days  
medianStepsPerday<-median(sumStepsPerday, na.rm=TRUE)
meanStepsPerday<-mean(sumStepsPerday, na.rm=TRUE)
```

**The median number of steps taken per day is `r medianStepsPerday`.**  
**The mean number of steps taken per day is `r meanStepsPerday`.**  

#What is the average daily activity pattern?  
```{r dailyActivity, echo=FALSE}
#remove intervals with NA steps
dX<-d[complete.cases(d),]
#calculate the mean steps in each interval
meanStepsPerInterval<-tapply(dX$steps, as.vector(dX$interval), mean)

plot(names(meanStepsPerInterval), meanStepsPerInterval, type = 'l', xlab="5-min Interval", ylab="Mean Number of Steps", main="Average Number of Steps Taken During the Day")
#find the interval with the max mean number of steps
maxSteps<-which.max(meanStepsPerInterval)

```
  
**The 5-min interval with the maximum number of steps is `r meanStepsPerInterval[maxSteps]`**  

#Input missing values
```{r missingSteps, echo=FALSE}
ok<-complete.cases(d)
notOK<-sum(!ok)

```
**The total number of missing values is `r notOK` ** 


```{r imputeMeanFunction, echo=FALSE}
#Fill in the missing step values (NA) with the mean for that 5-minute interval.  
#New dataset d2 contains the adjusted data.  
#-define function to replace NA with interval average  
#-impute.mean2 takes a dataframe as input
impute.mean <- function(x) {
    for (row in 1:nrow(x)){
        intervalLabel<-sprintf("%d",x$interval[row])
        if (is.na(x$steps[row])) {
            x$steps[row]<-meanStepsPerInterval[intervalLabel]
        }
    }
    x
}
```
 
```{r, echo=FALSE}
#-Create new dataframe with missing values replaced  
#-Calculate sum of steps taken per day  
#-Plot histogram  
#-Calculate median and mean steps taken per day 
d2<-impute.mean(d)

sumStepsPerday2<-tapply(d2$steps, as.vector(d2$date), sum)
hist(sumStepsPerday2, breaks = 20, ylab="Number of Days",
    xlab="Total Number of Steps",
    main="Histogram of Number of Adjusted Steps Taken Each Day
    Between Oct-01-2012 and Nov-30-2012")
medianStepsPerday2<-median(sumStepsPerday2, na.rm=TRUE)
meanStepsPerday2<-mean(sumStepsPerday2, na.rm=TRUE)
```
  
**With missing step values set as the average of the interval:**    
**the median number of steps taken per day is `r medianStepsPerday2`.**      
**the mean number of steps taken per day is `r meanStepsPerday2`.**      
```{r, echo=FALSE}
medianDiff<-medianStepsPerday - medianStepsPerday2
meanDiff<-meanStepsPerday - meanStepsPerday2
```
**The difference in median steps per day is**      
**`r medianDiff` for the original and adjusted data.**    
**The difference in mean number of steps per day is**    
**`r meanDiff` for the original and adjusted data.**    

#Are there differences in activity patterns between weekdays and weekends? 

```{r weekends, echo=FALSE}
#-Convert date string into date variable
d2$date<-as.Date(d2$date)
```

```{r, echo=FALSE}
#-Create new variable for weekend and weekdays
d2$weekend<-weekdays(d2$date) %in% c("Saturday", "Sunday")
d2$weekend<-ifelse( weekdays(d2$date) %in% c("Saturday", "Sunday"),"weekend", "weekday"  )
```
 
```{r, echo=FALSE}
#-make panel plots of time series plot of the 5-minute interval  
#and the average number of steps taken, averaged across all weekdays  
#or weekend days (y-axis) 
#Average steps per interval for weekdays and weekends
xx<-aggregate(steps ~ interval+weekend, data = d2, mean)
xyplot(steps~interval | weekend, type = 'l', data=xx,
    ylab="Number of steps", xlab="Interval", layout=c(1,2))

```